---
title: "Peer Assesment 1"
author: "K Samhith"
output:
  html_document:
    theme: journal
---
The complete project is documented here.

### Loading and preprocessing data

Download and unzip the data.

```{r}
if(!file.exists("getdata-projectfiles-UCI HAR Dataset.zip")) {
   temp <- tempfile()
  download.file("http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip",temp)
  unzip(temp)
  unlink(temp)
}


```

Now load the csv file into RStudio

```{r}
data <- read.csv("activity.csv")
```

###What is mean total number of steps taken per day?

Get rid of NA values

```{r}
data_nona<-na.omit(data)
```

Now aggregate the dataset and plot the histogram..!!

```{r}
data_daywise<-aggregate(steps~date,data_nona,sum)
hist(as.numeric(data_daywise$steps),col="red",main="Number of steps each day",xlab="Number of steps")


```

The mean and median are calculated

```{r}
mean_nona<-mean(data_daywise$steps)
median_nona<-median(data_daywise$steps)
```

The average is `r mean_nona` and the median is `r median_nona` .

###What is the average daily activity pattern?

```{r}

remove("mean")

 data_intwise<- aggregate(steps ~ interval, data_nona, mean)

plot(data_intwise$interval,data_intwise$steps, type="l", xlab="Interval", ylab="Number of steps",main="Average Number of Steps per Day by Interval")


```

The interval for maximum value is. 
```{r}
maxinterval<-data_intwise[data_intwise$steps==max(data_intwise$steps),1]
print(maxinterval)
```

###Imputing missing values

Inputting missing values with the average of their interval value.

```{r}
imputed_data<-data
for(i in 1:17568){
  if(is.na(imputed_data[i,1])){
    imputed_data[i,1]=data_intwise[data_intwise$interval==imputed_data[i,3],2]
  }
}  
```

Now arregating date wise.

```{r}
data3 <- aggregate(steps ~ date, imputed_data, sum)
```

Plotting the histogram.

```{r}
hist(data3$steps,col="red",xlab="Number of steps")
```

Calculating the new mean and median and calculating the difference with the old mean and median.

```{r}
newmean<-mean(data3$steps)
newmedian<-median(data3$steps)
diffmean<-mean_nona-newmean
diffmedian<-median_nona-newmedian
print(diffmean)
print(diffmedian)
```

###Are there differences in activity patterns between weekdays and weekends?

Converting the date column to date class and using weekday function to get the day and keeoing it in a new variable.

```{r}
imputed_data$date<-as.POSIXct(as.character(imputed_data$date),format="%Y-%m-%d")
imputed_data$day<-weekdays(imputed_data$date,abbreviate=FALSE)
```

Now converting it into two factor levels. Namely "weekday" and "weekend".

```{r}
for(i in 1:17568){
  if (imputed_data[i,4]=="Sunday"||imputed_data[i,4]=="Saturday"){
    imputed_data[i,4]="weekend"
  }
  else
    imputed_data[i,4]="weekday"
}


imputed_data$day<-as.factor(imputed_data$day)
```

Now making a panel plot.

```{r}
imputed_data$day<-as.factor(imputed_data$day)
data4<-aggregate(steps~interval+day,imputed_data,mean)
library(lattice)
xyplot(data4$steps~data4$interval|data4$day,layout=c(1,2),type="l",xlab="Intervals",ylab="Average number of steps")
```




